---
title: "BF Assignment 4"
output:
  html_document: default
  html_notebook: default
---


```{r}
data("austres")
class(austres)
data.ts <- austres
summary(data.ts)
head(data.ts)
data <- ts(austres,start=c(1971), frequency = 12)
```

```{r}
plot.ts(data)
```


The plot shows the increasing trend in the number of residents of Australia.


```{r}
ets_pred <- ets(data)
ets_pred
```

We can see that alpha = 6e-04 which means that prediction is done through average of whole dataset instead of through the resent value.

ETS(M,A,N) - 


Error type-"M"= Multiplicative,


Tread-None- "A"= Additive


Seasonality- None

```{r}
plot(ets_pred)
ets_pred$mse
```


Mean square error for the above forecast:  99.28186

Forecast:
Numbers (in thousands) of Australian residents measured quarterly from Jun 1978 to May 1979.

```{r}
forecast.ets(ets_pred, h=12)
```



Using function forecast ets() we can predict the lowest and highest value of  80% and 95% prediction interval of particular time interval.


```{r}
pred_ets <- forecast.ets(ets_pred, h=12)
plot(pred_ets)
```


In continuation of the above function forecast ets() , we can plot using function plot(pred_ets):
Forecasts for Jun(1978)-May(1979) marked as a Blue Line, for Dark Grey shaded area depicting 80% prediction, and for Light Grey shaded area depicting the 95% prediction interval.


