---
title: "BF Assignment 5"
output:
  html_document: default
  html_notebook: default
---


```{r}
data("AirPassengers")
class(AirPassengers)
AirPassengers
```



AirPassengers	Monthly Airline Passenger Numbers 1949-1960.


```{r}
plot.ts(AirPassengers)
```


We can check from the above Graph, the seasonal variations in the number of passengers per month.


```{r}
acf(AirPassengers, lag.max= 20)
```


From the above ACF graph we can see that all the lags are above the significant bound.


```{r}
stl_decomp <- stl(AirPassengers,s.window ="periodic")
stl_decomp
```


```{r}
plot(stl_decomp)
```


Plot Depiction: Top to bottom


Row1(Obeserved): Original Time series


Row2(trend): Trend component of time series and its increasing over the time.


Row3(seasonal): Seasonal variations in the time series


Row4(random): irregular or random component.


```{r}
attributes(stl_decomp)
```



```{r}
library(forecast)
stl_decomp_adj<-seasadj(stl_decomp)
stl_decomp_adj
plot(stl_decomp_adj)
lines(seasadj(stl_decomp))
```


By using the seasadj function, we have removed the seasonal component from the time series.


```{r}
f_stl <- forecast(stl_decomp,h=15)
f_stl
```


By using the forecast function, we are predicting the next 15 months from Jan 1961 to Mar 1962 with 80% and 95% confidence interval. For Mar 1962 we have range of(331.9300,589.4928) with 95% confidence.


```{r}
plot(f_stl)
```


Plot of forecast function, depicting 80% confidence( Dark Gray line) and 95% confidence(Light gray line).

Alternative Way:


```{r}
air_decompose<-decompose(AirPassengers)
air_decompose
attributes(air_decompose)
seasadj(air_decompose)
```


By using decompose function, we can see the time series components: Seasonal, Trend and Random.


```{r}
plot(air_decompose)
```


Plot Depiction: Top to bottom


Row1(Obeserved): Original Time series


Row2(trend): Trend component of time series and its increasing over the time.


Row3(seasonal): Seasonal variations in the time series


Row4(random): irregular or random component.




